DeepSphere: Efficient spherical Convolutional Neural Network with HEALPix sampling for cosmological applications

### 2018-10-25 scnn.tex
last todo + fix figures

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{eigenvectors}
	\caption{The first 16 eigenvectors of the graph Laplacian, an equivalent of Fourier modes, of a graph constructed from the HEALPix sampling of the sphere.
    Eigenvectors 1--3 could be associated with spherical harmonics of degree $\ell=1$ and order $|m|=(0,1)$, eigenvectors 4--8 with degree $\ell=2$ and order $|m|=(0,1,2)$, and eigenvectors 9--15 with degree $\ell=3$ and order $|m|=(0,1,2,3)$.
    Nevertheless, the graph eigenvectors are only approximating spherical harmonics.}
	\label{fig:graph_harmonics}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{filtering_speed}
    \caption{Comparison of filtering speed for Gaussian smoothing of maps of various sizes.
    The fast spherical harmonic transform (SHT) is implemented by the \texttt{healpy} Python package (via the \texttt{healpy.sphtfunc.smoothing} function).
    The graph filtering is defined by \eqnref{graph_convolution_cheby} and implemented with the \texttt{numpy} and \texttt{scipy} Python packages.
    Both are executed on a single core.
	The theoretical cost of filtering on the graph is $\bO(K N_{pix})$ and $\bO(\ell_{max}^3) = \bO(N_{pix}^{3/2})$ for the spherical harmonics, where $\ell_{max}$ is the largest angular frequency.}
    \label{fig:filtering_speed}
\end{figure}

@@

\begin{figure}
	\centering
    % \includegraphics[width=\linewidth]{pooling}
	\includegraphics[width=\linewidth]{figure_pooling_svg}
	\caption{Two levels of coarsening and pooling: groups of 4 cells are merged into one, then the data on them is summarized in one. The coarsest cell covers $1/12$ of the sphere.}
	\label{fig:pooling}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[height=7em]{sphere_grid_equirectangular}
    \hfill
    \includegraphics[height=7em]{sphere_grid_cubedsphere}
    \hfill
    \includegraphics[height=7em]{half_graph_4}
    \caption[]{Some pixelizations of the sphere. Left: the equirectangular grid, using equiangular spacing in a standard spherical-polar coordinate system. Middle: an equiangular cubed-sphere grid, as described in \citep{ronchi1996cubed}. Right: graph built from a HEALPix pixelization of half the sphere ($N_{side} = 4$). By construction, each node has eight neighbors, except the highlighted ones which have only seven.\footnotemark[4] Left and middle figures are taken from \citep{boomsma2017spherical}.}
    \label{fig:sphere_grids}
    \label{fig:healpix_graph_4}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{psd_sigma3}
	\caption{Power spectrum densities of the noiseless maps.
		To prevent the cosmological models to be distinguished by their power spectra alone, the maps have been smoothed with a Gaussian kernel of radius $3$ arcmins to remove high frequencies ($\ell>1000$).}
	\label{fig:psd_sigma3}
\end{figure}

@@

While testing the FCN and CNN variants, we made the following empirical observations.
First, training was more stable (in the sense that the loss decreased more smoothly) when using Chebyshev polynomials, as in \eqnref{graph_convolution_cheby}, rather than monomials, as in \eqnref{graph_convolution_monomial}.
Nevertheless, we could not observe a significant difference in accuracy after convergence.
Second, using $\ell_2$ regularization does not help either with performance or training of the models, mostly because the models are not over-fitting.
Third, we recommend initializing the Chebyshev coefficients with a centered Gaussian distribution with standard deviation $\sqrt{2/(F_{in} \times (K + 0.5))}$, where $K$ is the degree of the Chebyshev polynomials and $F_{in}$ the number of input channels. This standard deviation has the property to keep the energy of the signal more or less constant across layers.
\todo{Nath: cite the paper you read about that?}
% Nati to Michael: The problem with the previous claim is that it depends on the non-linarity. Furthermore, does it make sense to make such a claim if we use batch norm?
Finally, we observe that scaling the Laplacian's eigenvalues between $[-a, a]$, where $0 < a \leq 1$, significantly helps in stabilizing the optimization. We use $a = 0.75$ in our experiments.

@@

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{section_filter_last}
	\\
	\includegraphics[width=\linewidth]{gnonomic_filter_last}
	\caption{Random selection of 24 learned filters from the fifth spherical convolutional layer $L_5$. Top: section. Bottom: gnomonic projection.}
	\label{fig:learned_filter}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{graph_eigenvalues}
	\caption{The eigenvalues $\bLambda$ of the graph Laplacian $\L = \U \bLambda \U\trans$, which corresponds to squared frequencies, are clearly organized in groups. Each group corresponds to a degree $\ell$ of the spherical harmonics. Each degree has $2\ell + 1$ orders. See also \figref{graph_harmonics}.}
	\label{fig:graph_eigenvalues}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{subspace_harmonics_eigenvectors}
	\caption{Correspondence between the subspaces spanned by the graph Fourier modes and the spherical harmonics. Each pixel on the plot is the dot product between one spherical harmonic and one graph Fourier mode $\b u_i$. The plot clearly shows how the subspaces are aligned: modes and harmonics that correspond to different degrees $\ell$ are mostly orthogonal to each other, while modes and harmonics that correspond to the same degree $\ell$ are linearly dependent. There are $2 \ell + 1$ orders per degree $\ell$. Moreover, we see that the Fourier modes have the correct symmetries, indicated by the perfect orthogonality between even and odd functions. For example, the modes corresponding to $\ell=7$ are perfectly orthogonal to the harmonics corresponding to $\ell=0,2,4,6,8$.}
	\label{fig:subspace_harmonics_eigenvectors}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{border_effects}
	\caption{Convolution kernel (also called filter) $h$ localized in the center and left corner of a surface. A filter is localized on a pixel $i$ as $\T_i h = h(\L) \b \delta_i$ (see \secref{graph_convolution} and \eqnref{graph_convolution_spatial}). The graph is built from $1/12^\text{th}$ of the sphere at $N_{side} = 16$. The filter is not isotropic anymore when localized on the left corner due to border effects. The graph representation of a manifold assumes reflective border conditions.}
	\label{fig:border_effects}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{gaussian_filters_sphere}
  \caption{Comparison of convolution with the graph and the spherical harmonics ($N_{side} = 16$).
  Top: Diffusion of a unit of heat for different times $t$ using the graph.
  Bottom: spherical symmetric Gaussian smoothing for different $\sigma$ (arcmin).
  Relative difference between graph convolution and spherical smoothing: $10.4$\%, $4.8$\%, $3.8$\%.
}
  \label{fig:gaussian_filters_comparizon}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{gaussian_filters_spectral}
\caption{Visualization of the convolution kernel $K_t(x)=e^{-\tau t x}$ in spectral domain.}
\vspace{0.3in}
\includegraphics[width=\linewidth]{gaussian_filters_gnomonic}
\caption{Visualization of the convolution kernel $K_t(x)=e^{-\tau t x}$: cross-section of the kernel through the pixels along the equator.}
\vspace{0.3in}
\includegraphics[width=\linewidth]{gaussian_filters_section}
\caption{Visualization of the convolution kernel $K_t(x)=e^{-\tau t x}$ in a gnomonic projection.}
\label{fig:gaussian_filters_visualization}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{hist_error_order2_noise1_5}
	\caption{Error w.r.t.\ the number of training samples for histogram features and the linear SVM (setting: order $o=2$, relative noise level 1.5). The error clearly converged and adding more samples will not help. The validation error might be slightly below the training error as it is used to select hyper-parameters.}
	\label{fig:hist_error_evolution}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{sketch_lightcone_simulation}
	\caption{Left: Dark matter density field simulated using L-PICOLA for specific cosmological parameters. Right: Sketch of the lightcone constructed by arranging particles at different redshifts in concentric shells around the observer located at $z=0$.}
	\label{fig:sketch_lightcone_simulation}
\end{figure}

>>>>

\begin{figure}[t!]
	\centering
	\includegraphics[width=\linewidth]{eigenvectors}
	\caption{The first 16 eigenvectors of the graph Laplacian, an equivalent of Fourier modes, of a graph constructed from the HEALPix sampling of the sphere.
    Eigenvectors 1--3 could be associated with spherical harmonics of degree $\ell=1$ and order $|m|=(0,1)$, eigenvectors 4--8 with degree $\ell=2$ and order $|m|=(0,1,2)$, and eigenvectors 9--15 with degree $\ell=3$ and order $|m|=(0,1,2,3)$.
    Nevertheless, the graph eigenvectors are only approximating spherical harmonics.}
	\label{fig:graph_harmonics}
\end{figure}

@@

\begin{figure}[t!]
    \centering
    \includegraphics[width=\linewidth]{filtering_speed}
    \caption{Comparison of filtering speed for Gaussian smoothing of maps of various sizes.
    The fast spherical harmonic transform (SHT) is implemented by the \texttt{healpy} Python package (via the \texttt{healpy.sphtfunc.smoothing} function).
    The graph filtering is defined by \eqnref{graph_convolution_cheby} and implemented with the \texttt{numpy} and \texttt{scipy} Python packages.
    Both are executed on a single core.
	The theoretical cost of filtering on the graph is $\bO(K N_{pix})$ and $\bO(\ell_{max}^3) = \bO(N_{pix}^{3/2})$ for the spherical harmonics, where $\ell_{max}$ is the largest angular frequency.}
    \label{fig:filtering_speed}
\end{figure}

@@

\begin{figure}[t!]
	\centering
    % \includegraphics[width=\linewidth]{pooling}
	\includegraphics[width=\linewidth]{figure_pooling_svg}
	\caption{Two levels of coarsening and pooling: groups of 4 cells are merged into one, then the data on them is summarized in one. The coarsest cell covers $1/12$ of the sphere.}
	\label{fig:pooling}
\end{figure}

@@

\begin{figure}[t!]
    \centering
    \includegraphics[height=7em]{sphere_grid_equirectangular}
    \hfill
    \includegraphics[height=7em]{sphere_grid_cubedsphere}
    \hfill
    \includegraphics[height=7em]{half_graph_4}
    \caption[]{Some pixelizations of the sphere. Left: the equirectangular grid, using equiangular spacing in a standard spherical-polar coordinate system. Middle: an equiangular cubed-sphere grid, as described in \citep{ronchi1996cubed}. Right: graph built from a HEALPix pixelization of half the sphere ($N_{side} = 4$). By construction, each node has eight neighbors, except the highlighted ones which have only seven.\footnotemark[4] Left and middle figures are taken from \citep{boomsma2017spherical}.}
    \label{fig:sphere_grids}
    \label{fig:healpix_graph_4}
\end{figure}

@@

\begin{figure}[t!]
	\centering
	\includegraphics[width=\linewidth]{psd_sigma3}
	\caption{Power spectrum densities of the noiseless maps.
		To prevent the cosmological models to be distinguished by their power spectra alone, the maps have been smoothed with a Gaussian kernel of radius $3$ arcmins to remove high frequencies ($\ell>1000$).}
	\label{fig:psd_sigma3}
\end{figure}

@@

While testing the FCN and CNN variants, we made the following empirical observations.
First, training was more stable (in the sense that the loss decreased more smoothly) when using Chebyshev polynomials, as in \eqnref{graph_convolution_cheby}, rather than monomials, as in \eqnref{graph_convolution_monomial}.
Nevertheless, we could not observe a significant difference in accuracy after convergence.
Second, using $\ell_2$ regularization does not help either with performance or training of the models, mostly because the models are not over-fitting.
Third, we recommend initializing the Chebyshev coefficients with a centered Gaussian distribution with standard deviation $\sqrt{2/(F_{in} \times (K + 0.5))}$, where $K$ is the degree of the Chebyshev polynomials and $F_{in}$ the number of input channels. This standard deviation has the property to keep the energy of the signal more or less constant across layers\footnote{We derived this rule from the Chebyshev polynomial equations and some empirical experiments.}.
% Nati to Michael: The problem with the previous claim is that it depends on the non-linarity. Furthermore, does it make sense to make such a claim if we use batch norm?
Finally, we observe that scaling the Laplacian's eigenvalues between $[-a, a]$, where $0 < a \leq 1$, significantly helps in stabilizing the optimization. We use $a = 0.75$ in our experiments.

\begin{figure}
	\centering
	\includegraphics[width=0.85\linewidth]{section_filter_last}
	\vspace{0.5cm}\\
	\includegraphics[width=0.85\linewidth]{gnonomic_filter_last}
	\caption{Random selection of 24 learned filters from the fifth spherical convolutional layer $L_5$. Top: section. Bottom: gnomonic projection.}
	\label{fig:learned_filter}
\end{figure}

@@

@@

\begin{figure}[ht!]
	\centering
	\includegraphics[width=\linewidth]{graph_eigenvalues}
	\caption{The eigenvalues $\bLambda$ of the graph Laplacian $\L = \U \bLambda \U\trans$, which corresponds to squared frequencies, are clearly organized in groups. Each group corresponds to a degree $\ell$ of the spherical harmonics. Each degree has $2\ell + 1$ orders. See also \figref{graph_harmonics}.}
	\label{fig:graph_eigenvalues}
\end{figure}

\begin{figure}[ht!]
	\centering
	\includegraphics[width=\linewidth]{subspace_harmonics_eigenvectors}
	\caption{Correspondence between the subspaces spanned by the graph Fourier modes and the spherical harmonics. Each pixel on the plot is the dot product between one spherical harmonic and one graph Fourier mode $\b u_i$. The plot clearly shows how the subspaces are aligned: modes and harmonics that correspond to different degrees $\ell$ are mostly orthogonal to each other, while modes and harmonics that correspond to the same degree $\ell$ are linearly dependent. There are $2 \ell + 1$ orders per degree $\ell$. Moreover, we see that the Fourier modes have the correct symmetries, indicated by the perfect orthogonality between even and odd functions. For example, the modes corresponding to $\ell=7$ are perfectly orthogonal to the harmonics corresponding to $\ell=0,2,4,6,8$.}
	\label{fig:subspace_harmonics_eigenvectors}
\end{figure}

@@

\begin{figure}[ht!]
	\centering
	\includegraphics[width=\linewidth]{border_effects}
	\caption{Convolution kernel (also called filter) $h$ localized in the center and left corner of a surface. A filter is localized on a pixel $i$ as $\T_i h = h(\L) \b \delta_i$ (see \secref{graph_convolution} and \eqnref{graph_convolution_spatial}). The graph is built from $1/12^\text{th}$ of the sphere at $N_{side} = 16$. The filter is not isotropic anymore when localized on the left corner due to border effects. The graph representation of a manifold assumes reflective border conditions.}
	\label{fig:border_effects}
\end{figure}

@@

\begin{figure}[ht!]
  \centering
  \includegraphics[width=\linewidth]{gaussian_filters_sphere}
  \caption{Comparison of convolution with the graph and the spherical harmonics ($N_{side} = 16$).
  Top: Diffusion of a unit of heat for different times $t$ using the graph.
  Bottom: spherical symmetric Gaussian smoothing for different $\sigma$ (arcmin).
  Relative difference between graph convolution and spherical smoothing: $10.4$\%, $4.8$\%, $3.8$\%.
}
  \label{fig:gaussian_filters_comparizon}
\end{figure}

@@

\begin{figure}[ht!]
\centering
\includegraphics[width=\linewidth]{gaussian_filters_spectral}
\caption{Visualization of the convolution kernel $K_t(x)=e^{-\tau t x}$ in spectral domain.}
\vspace{0.3in}
\includegraphics[width=\linewidth]{gaussian_filters_gnomonic}
\caption{Visualization of the convolution kernel $K_t(x)=e^{-\tau t x}$: cross-section of the kernel through the pixels along the equator.}
\vspace{0.3in}
\includegraphics[width=\linewidth]{gaussian_filters_section}
\caption{Visualization of the convolution kernel $K_t(x)=e^{-\tau t x}$ in a gnomonic projection.}
\label{fig:gaussian_filters_visualization}
\end{figure}

@@

\begin{figure}[ht!]
	\centering
	\includegraphics[width=\linewidth]{hist_error_order2_noise1_5}
	\caption{Error w.r.t.\ the number of training samples for histogram features and the linear SVM (setting: order $o=2$, relative noise level 1.5). The error clearly converged and adding more samples will not help. The validation error might be slightly below the training error as it is used to select hyper-parameters.}
	\label{fig:hist_error_evolution}
\end{figure}

@@

\begin{figure}[ht!]
	\centering
	\includegraphics[width=\linewidth]{sketch_lightcone_simulation}
	\caption{Left: Dark matter density field simulated using L-PICOLA for specific cosmological parameters. Right: Sketch of the lightcone constructed by arranging particles at different redshifts in concentric shells around the observer located at $z=0$.}
	\label{fig:sketch_lightcone_simulation}
\end{figure}
